Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2023-07-11 popl24/hyperprop.tex
checkpoint

As we've discussed in our section on Related Work
(\ref{section-related}), previous authors have noted connections
between the traditional formulation of the MPC security model- aka the
simulator security model- and hyperproperty-style formulations of
program trace behavior \cite{8429300,wysstar} including probabilistic
noninterference. This formulation of hyperproperties that are at least
sound with respect to simulator security bridges a conceptual gap
between foundational approaches to security, and also motivates and
supports automated analysis.

In this section, we formulate a new hyperproperty that we show is
sound for passive security in the real/ideal model, called
\emph{noninterference modulo output ($\NIMO$)}. We do not claim
completeness with respect to passive security, but soundness
guarantees that enforcement of $\NIMO$ in our analyses implies passive
security. In Section \ref{section-bruteforce} we show how $\NIMO$ can
be automatically enforced in a brute-force manner and in Section
\ref{section-composition} we show how $\NIMO$ can be refined to obtain
compositional properties supporting efficient, scalable enforcement of
complex protocols that incorporate modular subprotocols, using
both YGC and GMW circuits as examples where gates are
subprotocols. The latter methodology demonstrates a particular
advantage of our intensional formulation.

@@

Despite the intuitive resonance between noninterference
and passive security, since MPC
protocols implement an ideal functionality, they cannot generally satisfy
noninterference. We can easily demonstrate this by using
Example \ref{example-he} as a failure witness, where we'll
assume $C = \{0,2\}$ and $H = \{1,3\}$:
$$
(\progd(\prog_{\ref{example-he}}))({\outv \mapsto 1}|\{\{\sx{1}{1} \mapsto 0, \sx{2}{1} \mapsto 1, \sx{3}{1} \mapsto 0 \} \})
= 1
$$
whereas:
$$
(\progd(\prog_{\ref{example-he}}))({\outv \mapsto 1}|\{\{\sx{1}{1} \mapsto 1, \sx{2}{1} \mapsto 1, \sx{3}{1} \mapsto 0 \} \})
= 0
$$

@@

From an information flow perspective, we can say that a $\NIMO$
protocol does not release information beyond what is declassified
via the public output. From a real/ideal perspective, we can
observe that $\NIMO$ implies that the simulator can use
the protocol itself to reconstruct the distribution of adversarial
views. That is, given the output of the ideal functionality
$\idealf(\store_S)$, the simulator can randomly choose any
$\store'$ such that $\store' =_C \store$ and $\idealf(\store'_S) =
\idealf(\store_S)$ and run the protocol with initial memory $\store'$-
$\NIMO$ guarantees ideal reconstruction of the real-world corrupt views. 
 However, there is a subtlety here in that the adversary
must be able to tractably reconstruct this knowledge. This is not
generally true, in particular, if $\idealf$ is a 1-way
function. Define:
\begin{definition}[Ideal Knowledge]
  Given a functionality $\idealf$ and output value $v$, the associated
  \emph{ideal knowledge}, denoted $\ik(\idealf,v)$ is
  $
  \{ \store\ |\ \idealf(\store) = v \}
  $.
  We say that $\idealf$ is \emph{pre-imageable} iff $\ik(\idealf, v)$ for all
  $v$ can be computed tractably.
\end{definition}
Now we can show our main result, that noninterference modulo output,
together with the assumption that $\idealf$ is pre-imageable, implies
passive security of a protocol.
\begin{theorem}
  \label{theorem-nimo}
  Assume given pre-imageable $\idealf$ and a protocol $\prog$ that
  correctly implements $\idealf$.  If $\NIMO(\prog)$
  then $\prog$ is passive secure.
\end{theorem}

\begin{proof}
  Let $H$ and $C$ be an arbitrary partition of the federation and
  suppose $\prog$ satisfies noninterference modulo output. Let
  $\iov(\prog) = (V,S)$ and $\flips(\prog) = F$ with output view $x
  \in V$ and let $\store$ be an arbitrary element of $\mems(S \cup
  F)$. Then the distribution of adversarial views in the real world
  is, by definition,
  $\condd{\progd(\prog)}{V_C}{\store_S}$.

  The simulator is given both $\store_{S_C}$ and
  $\idealf(\store_S)$.  The simulation $\SIM_C(\store_{S_C},
  \idealf(\store_S))$ can be defined as follows. First, some $\store'
  \in \ik(\idealf,\idealf(\store))$ is randomly chosen such that
  $\store' =_C \store_S$, as is a random tape $\store'' \in
  \mems(F)$\footnote{The real/ideal model allows consultation of a
  ``Random Oracle''.}. Then, the run of $(\store' \cup \store'',
  \prog)$ is evaluated in simulation, yielding $(\store^{\SIM},\varnothing)$.
  The simulation returns $\store^{\SIM}_{V_C}$ as a result.

  Now, since the random tape is selected in simulation from the same distribution
  that we assume for the real world, after selection of $\store'$ the
  probability that any particular $\store^{\SIM}_{V_C}$ is returned is by definition
  $
  \progd(\prog)(\store^{\SIM}_{V_C} | \store')
  $.
  Furthermore, since we assume that $\prog$ correctly implements $\idealf$, this
  means
  $
  \store_S =_C \store' \wedge
     \condd{\progd(\prog)}{\{ \outv \}}{\store_S} =
     \condd{\progd(\prog)}{\{ \outv \}}{\store'}
  $,
  so by the definition of noninterference modulo output,
  any choice of $\store'$ yields the same distribution $\condd{\progd(\prog)}{V_C}{\store_S}$.
  Therefore by definition
  $
   \progd(\SIM_C(\store_{S_C},\idealf(\store))) = \condd{\progd(\prog)}{V_C}{\store_S}
  $.
\end{proof}

As for noninterference, we can show that $\NIMO$ is equivalent
to saying that a given program does not change the probability
of honest secrets conditioned on corrupt secrets and views
(including the output view), as compared to conditioning on
just the corrupt secrets and input. In addition to providing
insights about $\NIMO$, this form is convenient for
verification techniques. 
\begin{lemma}
  \label{lemma-nimo}
  $\NIMO(\prog)$ iff for all $H$ and $C$ with $|C| \le |H|$ and
  $\store \in \mems(S \cup V)$:
  $$\condd{\progd(\prog)}{S_H}{\store_{(S \cup \{\outv\})_C}} =
  \condd{\progd(\prog)}{S_H}{\store_{(S\cup V)_C}} $$
  where  $\iov(\prog) = (S,V)$.
\end{lemma}

>>>>

As we've discussed in our section on Related Work
(\ref{section-related}), previous authors have noted connections
between the traditional formulation of the MPC security model- aka the
simulator security model- and hyperproperty-style formulations of
program trace behavior \cite{8429300,wysstar} including probabilistic
noninterference. In this Section we show for to formulate extensional
properties relevant to MPC passive security in our language model by
direct appeal to program distributions. In particular,
\emph{noninterference modulo output ($\NIMO$)} is shown to be sound
for MPC passive security. In Section \ref{section-automation} we will
show how $\NIMO$ can be automatically enforced directly via
whole-program analysis, or by automatically verifying composable,
intensional properties of circuit components.

@@

Despite the intuitive resonance between noninterference
and passive security, since MPC
protocols implement an ideal functionality, they cannot generally satisfy
noninterference. We can easily demonstrate this by using
Example \ref{example-gmw-andcircuit} as a failure witness, where we'll
assume $C = \{0,2\}$ and $H = \{1\}$:
$$
(\progd(\prog_{\ref{example-gmw-andcircuit}}))({\outv \mapsto 1}|\{\{\sx{1}{s1} \mapsto 0, \sx{2}{s2} \mapsto 1 \} \})
= 0
$$
whereas:
$$
(\progd(\prog_{\ref{example-gmw-andcircuit}}))({\outv \mapsto 1}|\{\{\sx{1}{s1} \mapsto 1, \sx{2}{s2} \mapsto 1 \} \})
= 1
$$

@@

From an information flow perspective, we can say that a $\NIMO$
protocol does not release information beyond what is declassified
via the public output. From a real/ideal perspective, we can
observe that $\NIMO$ implies that the simulator can use
the protocol itself to reconstruct the distribution of adversarial
views. That is, given the output of the ideal functionality
$\idealf(\store_S)$, the simulator can randomly choose any
$\store'$ such that $\store' =_C \store$ and $\idealf(\store'_S) =
\idealf(\store_S)$ and run the protocol with initial memory $\store'$-
$\NIMO$ guarantees ideal reconstruction of the real-world corrupt views. 
 However, there is a subtlety here in that the adversary
must be able to tractably reconstruct this knowledge. This is not
generally true, in particular, if $\idealf$ is a 1-way
function. Define:
\begin{definition}[Ideal Knowledge]
  Given a functionality $\idealf$ and output value $v$, the associated
  \emph{ideal knowledge}, denoted $\ik(\idealf,v)$ is
  $
  \{ \store\ |\ \idealf(\store) = v \}
  $.
  We say that $\idealf$ is \emph{pre-imageable} iff $\ik(\idealf, v)$ for all
  $v$ can be computed tractably.
\end{definition}
Now we can show our main result, that noninterference modulo output,
together with the assumption that $\idealf$ is pre-imageable, implies
passive security of a protocol. The proof is provided in Appendix
\ref{section-proofs}, and follows by essentially the same argument as
for the analogous property of ``perfect passive NI security'' in
\cite{8429300}.  
\begin{restatable}{theorem}{nimosecure}
  \label{theorem-nimo}
  Assume given pre-imageable $\idealf$ and a protocol $\prog$ that
  correctly implements $\idealf$.  If $\NIMO(\prog)$
  then $\prog$ is passive secure.
\end{restatable}
As for noninterference, we can show that $\NIMO$ is equivalent
to saying that a given program does not change the probability
of honest secrets conditioned on corrupt secrets and views
(including the output view), as compared to conditioning on
just the corrupt secrets and input. In addition to providing
insights about $\NIMO$, this form is convenient for
verification techniques. 
\begin{lemma}
  \label{lemma-nimo}
  $\NIMO(\prog)$ iff for all $H$ and $C$ with $|C| \le |H|$ and
  $\store \in \mems(S \cup V)$:
  $$\condd{\progd(\prog)}{S_H}{\store_{(S \cup \{\outv\})_C}} =
  \condd{\progd(\prog)}{S_H}{\store_{(S\cup V)_C}} $$
  where  $\iov(\prog) = (S,V)$.
\end{lemma}

### 2023-07-11 popl24/proofs.tex
checkpoint

>>>>

\nimosecure*

\begin{proof}
  Let $H$ and $C$ be an arbitrary partition of the federation and
  suppose $\prog$ satisfies noninterference modulo output. Let
  $\iov(\prog) = (V,S)$ and $\flips(\prog) = F$ with output view $\outv
  \in V$ and let $\store$ be an arbitrary element of $\mems(S \cup
  F)$. Then the distribution of adversarial views in the real world
  is, by definition,
  $\condd{\progd(\prog)}{V_C}{\store_S}$.

  The simulator is given both $\store_{S_C}$ and
  $\idealf(\store_S)$.  The simulation $\SIM_C(\store_{S_C},
  \idealf(\store_S))$ can be defined as follows. First, some $\store'
  \in \ik(\idealf,\idealf(\store))$ is randomly chosen such that
  $\store' =_C \store_S$, as is a random tape $\store'' \in
  \mems(F)$\footnote{The real/ideal model allows consultation of a
  ``Random Oracle''.}. Then, the run of $(\store' \cup \store'',
  \prog)$ is evaluated in simulation, yielding $(\store^{\SIM},\varnothing)$.
  The simulation returns $\store^{\SIM}_{V_C}$ as a result.

  Now, since the random tape is selected in simulation from the same distribution
  that we assume for the real world, after selection of $\store'$ the
  probability that any particular $\store^{\SIM}_{V_C}$ is returned is by definition
  $
  \progd(\prog)(\store^{\SIM}_{V_C} | \store')
  $.
  Furthermore, since we assume that $\prog$ correctly implements $\idealf$, this
  means
  $
  \store_S =_C \store' \wedge
     \condd{\progd(\prog)}{\{ \outv \}}{\store_S} =
     \condd{\progd(\prog)}{\{ \outv \}}{\store'}
  $,
  so by the definition of noninterference modulo output,
  any choice of $\store'$ yields the same distribution $\condd{\progd(\prog)}{V_C}{\store_S}$.
  Therefore by definition
  $
   \progd(\SIM_C(\store_{S_C},\idealf(\store))) = \condd{\progd(\prog)}{V_C}{\store_S}
  $.
\end{proof}
